---
title: "ENV 790.30 - Time Series Analysis for Energy Data | Spring 2023"
subtitle: "Assignment 2 - Due date 02/03/23"
author: "Yuxiang Ren"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: console
---

## Submission Instructions

You should open the .rmd file corresponding to this assignment on
RStudio. The file is available on our class repository on Github.

Once you have the file open on your local machine the first thing you
will do is rename the file such that it includes your first and last
name (e.g., "LuanaLima_TSA_A02_Sp23.Rmd"). Then change "Student Name" on
line 4 with your name.

Then you will start working through the assignment by **creating code
and output** that answer each question. Be sure to use this assignment
document. Your report should contain the answer to each question and any
plots/tables you obtained (when applicable).

When you have completed the assignment, **Knit** the text and code into
a single PDF file. Submit this pdf using Sakai.

## R packages

R packages needed for this assignment:"forecast","tseries", and "dplyr".
Install these packages, if you haven't done yet. Do not forget to load
them before running your script, since they are NOT default packages.\\

```{r}
#Load/install required package here
library(forecast)
library(tseries)
library(dplyr)
library(ggplot2)
library(xlsx)
```

## Data set information

Consider the data provided in the spreadsheet
"Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx"
on our **Data** folder. The data comes from the US Energy Information
and Administration and corresponds to the December 2022 Monthly Energy
Review. The spreadsheet is ready to be used. You will also find a $.csv$
version of the data
"Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source-Edit.csv".
You may use the function $read.table()$ to import the $.csv$ data in R.
Or refer to the file "M2_ImportingData_CSV_XLSX.Rmd" in our Lessons
folder for functions that are better suited for importing the $.xlsx$.

```{r}
#Importing data set

A02_rawdata <- read.xlsx(file="./Data/Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx",header=FALSE,startRow = 13,sheetIndex=1) 
read_col_names <- read.xlsx(file="./Data/Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx",header=FALSE,startRow = 11,endRow=11,sheetIndex=1)

colnames(A02_rawdata) <- read_col_names
head(A02_rawdata)


```

## Question 1

You will work only with the following columns: Total Biomass Energy
Production, Total Renewable Energy Production, Hydroelectric Power
Consumption. Create a data frame structure with these three time series
only. Use the command head() to verify your data.

```{r Q1}
#Q1
df_biomP_renewP_hydroC <- data.frame("BiomP"=A02_rawdata$`Total Biomass Energy Production`, 
                                     "RenewP"=A02_rawdata$`Total Renewable Energy Production`, 
                                     "HydroC"=A02_rawdata$`Hydroelectric Power Consumption`)

head(df_biomP_renewP_hydroC)

#with time 
df_date_brh <- cbind(A02_rawdata$Month,df_biomP_renewP_hydroC)
names(df_date_brh) <- c("Time", "BiomP", "RenewP", "HydroC")
```

## Question 2

Transform your data frame in a time series object and specify the
starting point and frequency of the time series using the function ts().

```{r Q2}
ts_df_biomP_renewP_hydroC <- ts(df_biomP_renewP_hydroC[,1:3], frequency = 12, start = c(1973, 1))


```

## Question 3

Compute mean and standard deviation for these three series.

```{r Q3}
#Total biomass energy production 
mean_biomP <- mean(df_biomP_renewP_hydroC$BiomP)
sd_biomP <- sd(df_biomP_renewP_hydroC$BioP)

#Total renewable energy production 
mean_renewP <- mean(df_biomP_renewP_hydroC$RenewP)
sd_renewP <- sd(df_biomP_renewP_hydroC$RenewP)

#Hydroelectric power consumption
mean_hydroC <- mean(df_biomP_renewP_hydroC$HydroC)
sd_hydroC <- sd(df_biomP_renewP_hydroC$HydroC)


```

## Question 4

Display and interpret the time series plot for each of these variables.
Try to make your plot as informative as possible by writing titles,
labels, etc. For each plot add a horizontal line at the mean of each
series in a different color.

```{r Q4}
#date for label
date <- df_date_brh[590,1]

#Total biomass energy production
plot_bioP <- ggplot(df_date_brh, aes(x=Time, y=BiomP)) +
            geom_line(color="Green")+
  geom_hline(aes(yintercept=mean_biomP), colour= "black", linetype="dashed")+
  geom_text(aes(date, mean_biomP, label = "Mean", vjust= -1))+
  ggtitle("Total Biomass Energy Production")+ xlab("Year")+
  ylab("Energy Production (Trillion Btu)")
plot_bioP

#Total renewable energy production 
plot_renewP <- ggplot(df_date_brh, aes(x=Time, y=RenewP)) +
            geom_line(color="red")+
  geom_hline(aes(yintercept=mean_renewP), colour= "black", linetype="dashed")+
  geom_text(aes(date, mean_renewP, label = "Mean", vjust= -1))+
  ggtitle("Total renewable energy production")+ xlab("Year")+ 
  ylab("Energy Production (Trillion Btu)")
plot_renewP

#Hydroelectric power consumption
plot_hydroC <- ggplot(df_date_brh, aes(x=Time, y=HydroC)) +
            geom_line(color="blue")+
  geom_hline(aes(yintercept=mean_hydroC), colour= "black", linetype="dashed")+
  geom_text(aes(date, mean_hydroC, label = "Mean", vjust= -1))+
  ggtitle("Hydroelectric power consumption")+ xlab("Year")+ 
  ylab("Energy Consumption (Trillion Btu)")
plot_hydroC

```

> Answer:
>
> (1)The overall trend for total biomass energy production is
> increasing. The obvious upward changes are between 1975 and 1990, and
> from 2000 to 2017. It can also be observed that at the beginning of
> 2020, biomass energy production suddenly dropped sharply and has
> gradually increased in recent years. The figure shows seasonal trend.
> (2)The overall trend for total renewable energy production is
> increasing. The increasing trend has become noticeable since 2000. The
> figure shows strong seasonality. (3)The overall change trend of the
> total consumption of hydroelectric power consumption is a slight
> decline. It can be seen that most of the monthly energy consumption
> before 2000 is greater than that after 2000.The figure shows strong
> seasonality.

## Question 5

Compute the correlation between these three series. Are they
significantly correlated? Explain your answer.

```{r Q5}
#correlation between three series
cor_all <- cor(df_biomP_renewP_hydroC[, c(1,2,3)])
cor_all

# biomP and renewP
cor_br <- cor.test(df_biomP_renewP_hydroC$BiomP,df_biomP_renewP_hydroC$RenewP)
cor_br
# biomP and hydroC
cor_bh <- cor.test(df_biomP_renewP_hydroC$BiomP,df_biomP_renewP_hydroC$HydroC)
cor_bh
# renewP and hydroC 
cor_rh <- cor.test(df_biomP_renewP_hydroC$RenewP,df_biomP_renewP_hydroC$HydroC)
cor_rh
```

> Answer: (1) Biomass energy production and renewable energy production
> have a strong positive correlation, which coefficient is 0.919.
> Additionally, the p-value is less than 0.05, the result is
> significant. (2) Biomass energy production and hydroelectric power
> consumption have a low negative correlation. The correlation
> coefficient of these two variables is -0.300. And because the p-value
> is less than 0.05, the result is significant. (3) Renewable energy
> production and hydroelectric power consumption have negligible or weak
> correlation due to the coefficient being -0.100. The p-value of these
> two variables is less than 0.05. The result is significant.

## Question 6

Compute the autocorrelation function from lag 1 up to lag 40 for these
three variables. What can you say about these plots? Do the three of
them have the same behavior?

```{r Q6}
biomP_acf=acf(ts_df_biomP_renewP_hydroC[,1],lag.max=40, type="correlation", plot=TRUE)
renewP_acf=acf(ts_df_biomP_renewP_hydroC[,2],lag.max=40, type="correlation", plot=TRUE)
hydroC_acf=acf(ts_df_biomP_renewP_hydroC[,3],lag.max=40, type="correlation", plot=TRUE)

```

> Answer: Both biomass and renewable energy production have a high
> degree of autocorrelation between themselves and lagged versions.
> Although both biomass and renewable's ACF are declining with lags, all
> lags are significant. However, some lags' value in the ACF of
> hydroelectric power consumption is insignificant. The change in the
> autocorrelation value of lag shows that the data might have some
> seasonality.

## Question 7

Compute the partial autocorrelation function from lag 1 to lag 40 for
these three variables. How these plots differ from the ones in Q6?

```{r Q7}
biomP_pacf=pacf(ts_df_biomP_renewP_hydroC[,1],lag.max=40, plot=TRUE)
renewP_pacf=pacf(ts_df_biomP_renewP_hydroC[,2],lag.max=40, plot=TRUE)
hydroC_pacf=pacf(ts_df_biomP_renewP_hydroC[,3],lag.max=40, plot=TRUE)

```

> Answer: Most lags' value drops and becomes insignificant. Meanwhile,
> more significant negative values emerge.
